---
title: "M2_bootstrap"
author: "Yohei Kiguchi"
format: html
editor: visual
---

# 14 Bootstrap

## Basic Policy

-   Bootstrap from the entire POLYMOD dataset

-   Sample size of bootstrap $\to$ The full size POLYMOD data

-   Mean and Covariance Matrix $\to$ All from this full-from-full-size bootstrap

-   Variance might be too small $\to$ Deal with the adaptive approach

-   number of iteration $\to$ 10,000 times

-   Return to 02_M2 to visualize the result

## 1. Import libraries and necessary data

```{r}
library(tidyverse)
library(here)
library(tictoc)

poly_df = readRDS(here('data', 'processed', 'poly_cleaned.rds'))
```

## 2. Hyper Parameters

```{r}
seed_val = 100
n_boot = 10000
sample_num = length(unique(poly_df$part_id))
```

## 3. Bootstrap Algorithm

### 3-1. Calculate Summary Statistics

```{r}
summary_stats_m2 = function(df) {
  
  degree_data = df %>%
    group_by(part_id) %>%
    summarise(degree = n_distinct(cont_id, na.rm = TRUE), .groups = "drop")
  
  # 1. Mean Degree
  mean_degree = mean(degree_data$degree)
  # var_degree = var(degree_data$degree)
  
  total_edges = nrow(df)
  
  # 2. Prop MM / Prop FF
  prop_mm = mean(df$part_gender == 'M' & df$cnt_gender == 'M', na.rm = TRUE)
  prop_ff = mean(df$part_gender == 'F' & df$cnt_gender == 'F', na.rm = TRUE)
  
  # 3. Prop Age (Age Difference)
  age_diff = abs(df$part_age - df$cnt_age)
  prop_age_0_5 = mean(age_diff >= 0 & age_diff <= 5, na.rm = TRUE)
  prop_age_6_15 = mean(age_diff >= 6 & age_diff <= 15, na.rm = TRUE)
  prop_age_20_35 = mean(age_diff >= 20 & age_diff <= 35, na.rm = TRUE)
  prop_age_over_40 = mean(age_diff > 40, na.rm = TRUE)
  
  # 4. Prop School
  is_school_edge = (df$part_age >= 5 & df$part_age <= 20) & 
    (df$cnt_age >= 5 & df$cnt_age <= 20)
  prop_school = sum(is_school_edge, na.rm = TRUE) / total_edges
  
  # 5. Prop Tails (Instead of Variance)
  prop_degree_le_2 = mean(degree_data$degree <= 2)
  prop_degree_ge_25 = mean(degree_data$degree >= 25)
  
  return(c(
    "Log Mean Degree" = log(mean_degree),
    "Prop Degree <= 2" = prop_degree_le_2,
    "Prop Degree >= 25" = prop_degree_ge_25,
    "Prop MM" = prop_mm,
    "Prop FF" = prop_ff,
    "Prop Age 0-5" = prop_age_0_5,
    "Prop Age 6-15" = prop_age_6_15,
    "Prop Age 20-35" = prop_age_20_35,
    "Prop Age > 40" = prop_age_over_40,
    "Prop School" = prop_school
  ))
}

```

### 3-2. Bootstrap Implementation

```{r}
run_bootstrap_m2 = function(df, seed_val, n_boot, sample_num) {
  
  boot_results = matrix(NA, nrow = n_boot, ncol = 10)
  colnames(boot_results) = c("Log Mean Degree", "Prop Degree <= 2",
                             "Prop Degree >= 25",
                             "Prop MM", "Prop FF",
                             "Prop Age 0-5", "Prop Age 6-15",
                             "Prop Age 20-35", "Prop Age > 40",
                             "Prop School")
  
  set.seed(seed_val)
  
  part_ids = unique(df$part_id)
  
  for (i in 1 : n_boot) {
    # 1. Resample IDs
    sampled_ids = sample(part_ids, sample_num, replace = TRUE)
    
    # 2. Dataframe with only IDs
    ids_df = tibble(part_id = sampled_ids)
    
    # 3. Join with original data
    sampled_df = ids_df %>%
      left_join(df, by = 'part_id', relationship = 'many-to-many')
    
    boot_results[i, ] = summary_stats_m2(sampled_df)
    
    if(i %% 1000 == 0) cat(".")
  }
  cat('finished!\n')
  
  return(boot_results)
}
```

## 4. Run Bootstrap

```{r}
tic()
boot_full_full = run_bootstrap_m2(poly_df, seed_val, n_boot, sample_num)
toc()
```

## 5. Save the result

```{r}
boot_data = list()
boot_data$result = boot_full_full
boot_data$n_boot = n_boot
boot_data$seed_val = seed_val
boot_data$sample_num = sample_num

saveRDS(boot_data, 'results/M2_bootstrap_result_4_2.rds')
```
