---
title: "01_BM_bootstrap"
author: "Yohei Kiguchi"
format: html
editor: visual
---

# 01 Bootstrap

## Basic Policy

-   Bootstrap from the entire POLYMOD dataset

-   Sample size of bootstrap $\to$ The full size POLYMOD data

-   Mean and Covariance Matrix $\to$ All from this full-from-full-size bootstrap

-   Variance might be too small $\to$ Deal with the adaptive approach

-   number of iteration $\to$ 10,000 times

## 1. Import libraries and necessary data

```{r}
library(tidyverse)
library(here)

poly_df = readRDS(here('data', 'processed', 'poly_cleaned.rds'))
```

## 2. Hyper Parameters

```{r}
seed_val = 100
n_boot = 10000
sample_num = length(unique(poly_df$part_id))
```

## 3. Bootstrap Algorithm

### 3-1. Calculate Summary Statistics

```{r}
summary_stats = function(df) {
  
  degree_data = df %>%
    group_by(part_id) %>%
    summarise(degree = n_distinct(cont_id, na.rm = TRUE), .groups = "drop")
  
  mean_degree = mean(degree_data$degree)
  var_degree = var(degree_data$degree)
  
  mean_age_diff = df %>%
    summarise(diff = mean(abs(part_age - cnt_age), na.rm = TRUE)) %>%
    pull(diff)
  
  prop_same_gender = df %>%
    summarise(prop = mean(part_gender == cnt_gender, na.rm = TRUE)) %>%
    pull(prop)
  
  prop_mm = mean(df$part_gender == 'M' & df$cnt_gender == 'M', na.rm = TRUE)
  
  prop_ff = mean(df$part_gender == 'F' & df$cnt_gender == 'F', na.rm = TRUE)
  
  return(c(
    "Mean Degree" = mean_degree,
    "Log Variance" = log(var_degree),
    "Mean Age Diff" = mean_age_diff,
    "Prop MM" = prop_mm,
    "Prop FF" = prop_ff
    )
  )
}
```

### 3-2. Bootstrap Implementation

```{r}
run_bootstrap = function(df, seed_val, n_boot, sample_num) {
  
  boot_results = matrix(NA, nrow = n_boot, ncol = 5)
  colnames(boot_results) = c("Mean Degree", "Log Variance", "Mean Age Diff",
                             "Prop MM", "Prop FF")
  
  set.seed(seed_val)
  
  part_ids = unique(df$part_id)
  
  for (i in 1 : n_boot) {
    # 1. Resample IDs
    sampled_ids = sample(part_ids, sample_num, replace = TRUE)
    
    # 2. Dataframe with only IDs
    ids_df = tibble(part_id = sampled_ids)
    
    # 3. Join with original data
    sampled_df = ids_df %>%
      left_join(df, by = 'part_id', relationship = 'many-to-many')
    
    boot_results[i, ] = summary_stats(sampled_df)
    
    if(i %% 1000 == 0) cat(".")
  }
  cat('finished!\n')
  
  return(boot_results)
}
```

## 4. Run Bootstrap

```{r}
boot_full_full = run_bootstrap(poly_df, seed_val, n_boot, sample_num)
```

## 5. Save the result

```{r}
boot_data = list()
boot_data$result = boot_full_full
boot_data$n_boot = n_boot
boot_data$seed_val = seed_val
boot_data$sample_num = sample_num

saveRDS(boot_data, 'results/bootstrap_result.rds')
```
